---
title: "A New Development Index"
author: Michael Monzillo
subtitle: A Case Study in Data Visualization and Analysis in R
---

# Introduction
The UN Development Program every year publishes it's Human Development Index. The goal of this measure is to measure human development by country with respect to economic output, education level and health outcome. The goal of this project will be to reimagine such an index to account for political stability, economic stability, and economic sustainability. The index will be calculated and analyzed for nations comprising Latin America.

# Materials and methods

This section will describe data sources and coding methods. Before proceeding to the code and data itself however, it is important to discuss how this index will be defined.

Being that the Index is a model, it will be important to understand the key assumptions it
relies on, namely how it's inputs represent what they claim to represent. 

## We can consider these casewise:

1. Political Stability - This will be represented by the level of public (or government)
debt in a single country. The thinking here is that as public debt levels increase, a 
nation's government will become increasingly unstable.
2. Economic Stability - One of the most common macroeconomic indicators of economic 
stability is the inflation rate. The Consumer Price Index of each country, a measure
tracking the prices of all goods in a country intended for end consumers, will be included. An unchecked inflation rate is always a danger to any economy.
3. Environmental Sustainability - The are a plethora of measure that may be used to track a nation's environmental sustainability. Although some of these measures may change over time, levels of Carbon Dioxide Emissions are always a concern. Therefore CO2 emissions by country will be included.


Thus the general workflow proceeds as follows:

* Gathering Requisite Data
* Constructing the Index
* Examine Trends


Here's my first code chunk.
```{r}
1+2
```

Here's a quick look at the packages used for this project

```{r, message=F, warning=F}
library(fredr)
library(wbstats)
library(dplyr)
library(ggplot2)
library(foreach)
library(spData)
library(sf)
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```

## Download and clean all required data
```{r}
n=20
data=data.frame(x=runif(n,-180,180),
                y=runif(n,-60,60),
                size = runif(n, 5, 20),
                category = factor(
                  sample(letters[1:5], n, replace = TRUE)
                  ),
                value = rnorm(n))
```

```{r, results='asis'}
data %>% 
  slice(1:10) %>% #show only 1:n rows
  kable(digits=2,align="c")%>% #make table and round to two digits
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive")) #apply other formatting
```

Add any additional processing steps here.

# Results

Show tables, plots, etc. and describe them.

```{r, fig.width=6, fig.height=3, fig.cap="Map of completely random data"}
m <- leaflet(data) %>% 
  addTiles() %>% 
  addCircleMarkers(~x, ~y, radius = ~size,color = ~as.factor(category)) %>% 
  addPopups(~x[2], ~y[2], "Random popup")
m  # a map with the default OSM tile layer
```


```{r}
data %>% 
  ggplot(aes(x=x,y=y,col=category))+
  geom_point()
```

# Conclusions

What have you learned?  Are there any broader implications?

# References
