---
title: "A New Development Index"
author: Michael Monzillo
subtitle: A Case Study in Data Visualization and Analysis in R
---

# Introduction
The UN Development Program every year publishes it's Human Development Index. The goal of this measure is to measure human development by country with respect to economic output, education level and health outcome. The goal of this project will be to reimagine such an index to account for political stability, economic stability, and economic sustainability. The index will be calculated and analyzed for nations comprising Latin America.
```{r include=FALSE}
library(fredr)
library(wbstats)
library(dplyr)
library(ggplot2)
library(foreach)
library(spData)
library(sf)
library(plm)
library(forecast)
library(rworldmap)
library(ggmap)


fredr_set_key("6b0cdd4cca87cf2616946fd1b9c3e865")

# Political Stability - Indicator = Public Debt



Arg = fredr_series_search_text("Argentina debt GDP")
Arg$title
Arg$id[5]
Bol = fredr_series_search_text("Bolivia public debt")
Bol$title
Bol$id[3]
Bra = fredr_series_search_text("Brazil public debt")
Bra$title
Bra$id[5]
Chi = fredr_series_search_text("Chile public debt")
Chi$title
Chi$id[2]
Col = fredr_series_search_text("Colombia public debt")
Col$title
Col$id[7]
Cos = fredr_series_search_text("Costa Rica public debt")
Cos$title
Cos$id[3]
Dom = fredr_series_search_text("Dominican Republic public debt")
Dom$title
Dom$id[6]
Ecu = fredr_series_search_text("Ecuador public debt")
Ecu$title
Ecu$id[4]
Els = fredr_series_search_text("El Salvador public debt")
Els$title
Els$id[3]
Mex = fredr_series_search_text("Mexico public debt")
Mex$title
Mex$id[5]
Pan = fredr_series_search_text("Panama public debt")
Pan$title
Pan$id[4]
Par = fredr_series_search_text("Paraguay public debt")
Par$title
Par$id[3]
Per = fredr_series_search_text("Peru public debt")
Per$title
Per$id[7]
Uru = fredr_series_search_text("Uruguay public debt")
Uru$title
Uru$id[5]


dates = c(1999:2014)
dates.1 = c(1980:2017)
debtnames = c("date", "debt")


Arg.debt = fredr("DDDM07ARA156NWDB")
Arg.debt
Arg.debt.1 = cbind(dates.1, Arg.debt) %>%
  filter(dates.1 >= 2000) %>%
  filter(dates.1 <= 2014) %>%
  dplyr::select(dates.1, value) %>%
  arrange((desc(dates.1)))
Arg.debt.1
colnames(Arg.debt.1) <- c("date", "Arg.debt")
Arg.debt.1

Bol.debt = fredr("DDDM07BOA156NWDB")
head(Bol.debt)
tail(Bol.debt)
Bol.debt.1 = cbind(dates.1, Bol.debt) %>%
  filter(dates.1 >= 2000) %>%
  filter(dates.1 <= 2014) %>%
  dplyr::select(dates.1, value) %>%
  arrange((desc(dates.1))) 
Bol.debt.1
colnames(Bol.debt.1) <- c("date", "Bol.debt")
Bol.debt.1

Bra.debt = fredr("DDDM07BRA156NWDB")
head(Bra.debt)
tail(Bra.debt)
dates.2 = c(1982:2017)
Bra.debt.1 = cbind(dates.2, Bra.debt) %>%
  filter(dates.2 >= 2000) %>%
  filter(dates.2 <= 2014) %>%
  dplyr::select(dates.2, value)  %>%
  arrange((desc(dates.2)))
Bra.debt.1
colnames(Bra.debt.1) <- c("date", "Bra.debt")
Bra.debt.1

Chi.debt = fredr("DDDM07CLA156NWDB")
head(Chi.debt)
tail(Chi.debt)
Chi.debt.1 = cbind(dates.1, Chi.debt) %>%
  filter(dates.1 >= 2000) %>%
  filter(dates.1 <= 2014) %>%
  dplyr::select(dates.1, value)  %>%
  arrange((desc(dates.1)))
Chi.debt.1
colnames(Chi.debt.1) <- c("date", "Chi.debt")
Chi.debt.1

Col.debt = fredr("DDDM07COA156NWDB")
head(Col.debt)
tail(Col.debt)
Col.debt.1 = cbind(dates.1, Col.debt) %>%
  filter(dates.1 >= 2000) %>%
  filter(dates.1 <= 2014) %>%
  dplyr::select(dates.1, value)  %>%
  arrange((desc(dates.1)))
Col.debt.1
colnames(Col.debt.1) <- c("date", "Col.debt")
Col.debt.1

Cos.debt = fredr("DDDM04COA156NWDB")
head(Cos.debt)
tail(Cos.debt)
dates.3=c(1989:2017)
Cos.debt.1 = cbind(dates.3, Cos.debt) %>%
  filter(dates.3 >= 2000) %>%
  filter(dates.3 <= 2014) %>%
  dplyr::select(dates.3, value)  %>%
  arrange((desc(dates.3)))
Cos.debt.1
colnames(Cos.debt.1) <- c("date", "Cos.debt")
Cos.debt.1

Dom.debt = fredr("DDDM07DOA156NWDB")
head(Dom.debt)
tail(Dom.debt)
dates.4 = c(1994:2017)
Dom.debt.1 = cbind(dates.4, Dom.debt) %>%
  filter(dates.4 >= 2000) %>%
  filter(dates.4 <= 2014) %>%
  dplyr::select(dates.4, value)  %>%
  arrange((desc(dates.4)))
Dom.debt.1
colnames(Dom.debt.1) <- c("date", "Dom.debt")
Dom.debt.1

Ecu.debt = fredr("DDDM07ECA156NWDB")
head(Ecu.debt)
tail(Ecu.debt)
Ecu.debt.1 = cbind(dates.1, Ecu.debt) %>%
  filter(dates.1 >= 2000) %>%
  filter(dates.1 <= 2014) %>%
  dplyr::select(dates.1, value)  %>%
  arrange((desc(dates.1)))
Ecu.debt.1
colnames(Ecu.debt.1) <- c("date", "Ecu.debt")
Ecu.debt.1

Els.debt = fredr("DDDM07SVA156NWDB")
head(Els.debt)  
tail(Els.debt)
dates.9 = c(1997:2017)
Els.debt.1 = cbind(dates.9, Els.debt) %>%
  filter(dates.9 >= 2000) %>%
  filter(dates.9 <= 2014) %>%
  dplyr::select(dates.9, value)  %>%
  arrange((desc(dates.9)))
Els.debt.1
colnames(Els.debt.1) <- c("date", "Els.debt")
Els.debt.1


Mex.debt = fredr("DDDM07MXA156NWDB")
head(Mex.debt)
tail(Mex.debt)
Mex.debt.1 = cbind(dates.1, Mex.debt) %>%
  filter(dates.1 >= 2000) %>%
  filter(dates.1 <= 2014) %>%
  dplyr::select(dates.1, value)  %>%
  arrange((desc(dates.1)))
Mex.debt.1
colnames(Mex.debt.1) <- c("date", "Mex.debt")
Mex.debt.1

Pan.debt = fredr("DDDM07PAA156NWDB")
head(Pan.debt)
tail(Pan.debt)
Pan.debt.1 = cbind(dates.1, Pan.debt) %>%
  filter(dates.1 >= 2000) %>%
  filter(dates.1 <= 2014) %>%
  dplyr::select(dates.1, value)  %>%
  arrange((desc(dates.1)))
Pan.debt.1
colnames(Pan.debt.1) <- c("date", "Pan.debt")
Pan.debt.1

Par.debt = fredr("DDDM10PYA156NWDB")
head(Par.debt)
tail(Par.debt)
dates.6 = c(2000:2017)
Par.debt.1 = cbind(dates.6, Par.debt) %>%
  filter(dates.6 >= 2000) %>%
  filter(dates.6 <= 2014) %>%
  dplyr::select(dates.6, value)  %>%
  arrange((desc(dates.6)))
Par.debt.1
colnames(Par.debt.1) <- c("date", "Par.debt")
Par.debt.1

Per.debt = fredr("DDDM07PEA156NWDB")
head(Per.debt)
tail(Per.debt)
Per.debt.1 = cbind(dates.4, Per.debt) %>%
  filter(dates.4 >= 2000) %>%
  filter(dates.4 <= 2014) %>%
  dplyr::select(dates.4, value)  %>%
  arrange((desc(dates.4)))
Per.debt.1
colnames(Per.debt.1) <- c("date", "Per.debt")
Per.debt.1

Uru.debt = fredr("DDDM07UYA156NWDB")
head(Uru.debt)
tail(Uru.debt)
dates.7 = c(1991:2017)
Uru.debt.1 = cbind(dates.7, Uru.debt) %>%
  filter(dates.7 >= 2000) %>%
  filter(dates.7 <= 2014) %>%
  dplyr::select(dates.7, value)  %>%
  arrange((desc(dates.7)))
Uru.debt.1
colnames(Uru.debt.1) <- c("date", "Uru.debt")
Uru.debt.1


# Environmental Sustainability - Indicator = CO2 Emissions

wbsearch(pattern = " CO2 emission")

CO2 = wb(indicator = "EN.ATM.CO2E.KT")

Arg.CO2 = CO2 %>%
  filter(country == "Argentina") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value) %>%
  arrange((desc(date)))
colnames(Arg.CO2) <- c("date1", "Arg.CO2")
Arg.CO2


Bol.CO2 = CO2 %>%
  filter(country == "Bolivia") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Bol.CO2) <- c("date1", "Bol.CO2")
Bol.CO2

Bra.CO2 = CO2 %>%
  filter(country == "Brazil") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Bra.CO2) <- c("date1", "Bra.CO2")
Bra.CO2

Chi.CO2 = CO2 %>%
  filter(country == "Chile") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Chi.CO2) <- c("date1", "Chi.CO2")
Chi.CO2

Col.CO2 = CO2 %>%
  filter(country == "Colombia") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Col.CO2) <- c("date1", "Col.CO2")
Col.CO2

Cos.CO2 = CO2 %>%
  filter(country == "Costa Rica") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Cos.CO2) <- c("date1", "Cos.CO2")
Cos.CO2

Dom.CO2 = CO2 %>%
  filter(country == "Dominican Republic") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Dom.CO2) <- c("date1", "Dom.CO2")
Dom.CO2

Ecu.CO2 = CO2 %>%
  filter(country == "Ecuador") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Ecu.CO2) <- c("date1", "Ecu.CO2")
Ecu.CO2

Els.CO2 = CO2 %>%
  filter(country == "El Salvador") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Els.CO2) <- c("date1", "Els.CO2")
Els.CO2

Mex.CO2 = CO2 %>%
  filter(country == "Mexico") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Mex.CO2) <- c("date1", "Mex.CO2")
Mex.CO2

Pan.CO2 = CO2 %>%
  filter(country == "Panama") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Pan.CO2) <- c("date1", "Pan.CO2")
Pan.CO2

Par.CO2 = CO2 %>%
  filter(country == "Paraguay") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Par.CO2) <- c("date1", "Par.CO2")
Par.CO2

Per.CO2 = CO2 %>%
  filter(country == "Peru") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Per.CO2) <- c("date1", "Per.CO2")
Per.CO2

Uru.CO2 = CO2 %>%
  filter(country == "Uruguay") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value)
colnames(Uru.CO2) <- c("date1", "Uru.CO2")
Uru.CO2

# Economic Stability = Inflation Rate

Arg.CPI = fredr_series_search_text("Argentina CPI")
Arg.CPI$title
Arg.CPI$id[2]
Arg.CPI = fredr("DDOE01ARA086NWDB")
head(Arg.CPI)
tail(Arg.CPI)
CPIdates.1 = c(1960:2014)
Arg.CPI.1 = cbind(CPIdates.1, Arg.CPI) %>%
  filter(CPIdates.1 >= 2000) %>%
  filter(CPIdates.1 <= 2014) %>%
  dplyr::select(CPIdates.1, value) %>%
  arrange((desc(CPIdates.1)))
Arg.CPI.1
colnames(Arg.CPI.1) <- c("date2", "Arg.CPI")
Arg.CPI.1

Bol = fredr_series_search_text("Bolivia CPI")
Bol$title
Bol$id[2]
Bol.CPI = fredr("DDOE01BOA086NWDB")
head(Bol.CPI)
tail(Bol.CPI)
CPIdates.2 = c(1960:2017)
Bol.CPI.1 = cbind(CPIdates.2, Bol.CPI) %>%
  filter(CPIdates.2 >= 2000) %>%
  filter(CPIdates.2 <= 2014) %>%
  dplyr::select(CPIdates.2, value) %>%
  arrange((desc(CPIdates.2)))
Bol.CPI.1
colnames(Bol.CPI.1) <- c("date2", "Bol.CPI")
Bol.CPI.1

Bra = fredr_series_search_text("Brazil CPI")
Bra$title
Bra$id[16]
Bra.CPI = fredr("DDOE01BRA086NWDB")
head(Bra.CPI)
tail(Bra.CPI)
CPIdates.3 = c(1979:2017)
Bra.CPI.1 = cbind(CPIdates.3, Bra.CPI) %>%
  filter(CPIdates.3 >= 2000) %>%
  filter(CPIdates.3 <= 2014) %>%
  dplyr::select(CPIdates.3, value) %>%
  arrange((desc(CPIdates.3)))
Bra.CPI.1
colnames(Bra.CPI.1) <- c("date2", "Bra.CPI")
Bra.CPI.1

Chi = fredr_series_search_text("Chile CPI")
Chi$title
Chi$id[5]
Chi.CPI = fredr("DDOE01CLA086NWDB")
head(Chi.CPI)
tail(Chi.CPI)
Chi.CPI.1 = cbind(CPIdates.2, Chi.CPI) %>%
  filter(CPIdates.2 >= 2000) %>%
  filter(CPIdates.2 <= 2014) %>%
  dplyr::select(CPIdates.2, value) %>%
  arrange((desc(CPIdates.2)))
Chi.CPI.1
colnames(Chi.CPI.1) <- c("date2", "Chi.CPI")
Chi.CPI.1

Col = fredr_series_search_text("Colombia CPI")
Col$title
Col$id[1]
Col.CPI = fredr("DDOE01COA086NWDB")
head(Col.CPI)
tail(Col.CPI)
Col.CPI.1 = cbind(CPIdates.2, Col.CPI) %>%
  filter(CPIdates.2 >= 2000) %>%
  filter(CPIdates.2 <= 2014) %>%
  dplyr::select(CPIdates.2, value) %>%
  arrange((desc(CPIdates.2)))
Col.CPI.1
colnames(Col.CPI.1) <- c("date2", "Col.CPI")
Col.CPI.1

Cos = fredr_series_search_text("Costa Rica CPI")
Cos$title
Cos$id[2]
Cos.CPI = fredr("DDOE02CRA086NWDB")
head(Cos.CPI)
tail(Cos.CPI)
Cos.CPI.1 = cbind(CPIdates.2, Cos.CPI) %>%
  filter(CPIdates.2 >= 2000) %>%
  filter(CPIdates.2 <= 2014) %>%
  dplyr::select(CPIdates.2, value) %>%
  arrange((desc(CPIdates.2)))
Cos.CPI.1
colnames(Cos.CPI.1) <- c("date2", "Cos.CPI")
Cos.CPI.1

Dom = fredr_series_search_text("Dominican Republic CPI")
Dom$title
Dom$id[2]
Dom.CPI = fredr("DDOE01DOA086NWDB")
head(Dom.CPI)
tail(Dom.CPI)
Dom.CPI.1 = cbind(CPIdates.2, Dom.CPI) %>%
  filter(CPIdates.2 >= 2000) %>%
  filter(CPIdates.2 <= 2014) %>%
  dplyr::select(CPIdates.2, value) %>%
  arrange((desc(CPIdates.2)))
Dom.CPI.1
colnames(Dom.CPI.1) <- c("date2", "Dom.CPI")
Dom.CPI.1

Ecu = fredr_series_search_text("Ecuador CPI")
Ecu$title
Ecu$id[2]
Ecu.CPI = fredr("DDOE01ECA086NWDB")
head(Ecu.CPI)
tail(Ecu.CPI)
Ecu.CPI.1 = cbind(CPIdates.2, Ecu.CPI) %>%
  filter(CPIdates.2 >= 2000) %>%
  filter(CPIdates.2 <= 2014) %>%
  dplyr::select(CPIdates.2, value) %>%
  arrange((desc(CPIdates.2)))
Ecu.CPI.1
colnames(Ecu.CPI.1) <- c("date2", "Ecu.CPI")
Ecu.CPI.1

Els = fredr_series_search_text("El Salvador CPI")
Els$title
Els$id[2]
Els.CPI = fredr("DDOE02SVA086NWDB")
head(Els.CPI)
tail(Els.CPI)
Els.CPI.1 = cbind(CPIdates.2, Els.CPI) %>%
  filter(CPIdates.2 >= 2000) %>%
  filter(CPIdates.2 <= 2014) %>%
  dplyr::select(CPIdates.2, value) %>%
  arrange((desc(CPIdates.2)))
Els.CPI.1
colnames(Els.CPI.1) <- c("date2", "Els.CPI")
Els.CPI.1


Mex = fredr_series_search_text("Mexico CPI")
Mex$title
Mex$id[29]
Mex.CPI = fredr("DDOE02MXA086NWDB")
head(Mex.CPI)
tail(Mex.CPI)
Mex.CPI.1 = cbind(CPIdates.2, Mex.CPI) %>%
  filter(CPIdates.2 >= 2000) %>%
  filter(CPIdates.2 <= 2014) %>%
  dplyr::select(CPIdates.2, value) %>%
  arrange((desc(CPIdates.2)))
Mex.CPI.1
colnames(Mex.CPI.1) <- c("date2", "Mex.CPI")
Mex.CPI.1

Pan = fredr_series_search_text("Panama CPI")
Pan$title
Pan$id[1]
Pan.CPI = fredr("DDOE02PAA086NWDB")
head(Pan.CPI)
tail(Pan.CPI)
Pan.CPI.1 = cbind(CPIdates.2, Pan.CPI) %>%
  filter(CPIdates.2 >= 2000) %>%
  filter(CPIdates.2 <= 2014) %>%
  dplyr::select(CPIdates.2, value) %>%
  arrange((desc(CPIdates.2)))
Pan.CPI.1
colnames(Pan.CPI.1) <- c("date2", "Pan.CPI")
Pan.CPI.1

Par = fredr_series_search_text("Paraguay CPI")
Par$title
Par$id[2]
Par.CPI = fredr("DDOE01PYA086NWDB")
head(Par.CPI)
tail(Par.CPI)
Par.CPI.1 = cbind(CPIdates.2, Par.CPI) %>%
  filter(CPIdates.2 >= 2000) %>%
  filter(CPIdates.2 <= 2014) %>%
  dplyr::select(CPIdates.2, value) %>%
  arrange((desc(CPIdates.2)))
Par.CPI.1
colnames(Par.CPI.1) <- c("date2", "Par.CPI")
Par.CPI.1

Per = fredr_series_search_text("Peru CPI")
Per$title
Per$id[2]
Per.CPI = fredr("DDOE01PEA086NWDB")
head(Per.CPI)
tail(Per.CPI)
Per.CPI.1 = cbind(CPIdates.2, Per.CPI) %>%
  filter(CPIdates.2 >= 2000) %>%
  filter(CPIdates.2 <= 2014) %>%
  dplyr::select(CPIdates.2, value) %>%
  arrange((desc(CPIdates.2)))
Per.CPI.1
colnames(Per.CPI.1) <- c("date2", "Per.CPI")
Per.CPI.1

Uru = fredr_series_search_text("Uruguay CPI")
Uru$title
Uru$id[1]
Uru.CPI = fredr("DDOE01UYA086NWDB")
head(Uru.CPI)
tail(Uru.CPI)
Uru.CPI.1 = cbind(CPIdates.2, Uru.CPI) %>%
  filter(CPIdates.2 >= 2000) %>%
  filter(CPIdates.2 <= 2014) %>%
  dplyr::select(CPIdates.2, value) %>%
  arrange((desc(CPIdates.2)))
Uru.CPI.1
colnames(Uru.CPI.1) <- c("date2", "Uru.CPI")
Uru.CPI.1

# Create Country Dataframes, Gather Geometric Means

Arg = cbind(Arg.debt.1, Arg.CO2) 
Arg = cbind(Arg, Arg.CPI.1) %>%
  dplyr::select(date, Arg.debt, Arg.CO2, Arg.CPI)
Arg.mean= ((Arg$Arg.debt*Arg$Arg.CO2*Arg$Arg.CPI)^(1/3))
Arg = cbind(Arg, Arg.mean)
Arg

Bol = cbind(Bol.debt.1, Bol.CO2) 
Bol = cbind(Bol, Bol.CPI.1) %>%
  dplyr::select(date, Bol.debt, Bol.CO2, Bol.CPI)
Bol.mean= ((Bol$Bol.debt*Bol$Bol.CO2*Bol$Bol.CPI)^(1/3))
Bol = cbind(Bol, Bol.mean)
Bol

Bra = cbind(Bra.debt.1, Bra.CO2) 
Bra = cbind(Bra, Bra.CPI.1) %>%
  dplyr::select(date, Bra.debt, Bra.CO2, Bra.CPI)
Bra.mean= ((Bra$Bra.debt*Bra$Bra.CO2*Bra$Bra.CPI)^(1/3))
Bra = cbind(Bra, Bra.mean)
Bra

Chi = cbind(Chi.debt.1, Chi.CO2) 
Chi = cbind(Chi, Chi.CPI.1) %>%
  dplyr::select(date, Chi.debt, Chi.CO2, Chi.CPI)
Chi.mean= ((Chi$Chi.debt*Chi$Chi.CO2*Chi$Chi.CPI)^(1/3))
Chi = cbind(Chi, Chi.mean)
Chi

Col = cbind(Col.debt.1, Col.CO2) 
Col = cbind(Col, Col.CPI.1) %>%
  dplyr::select(date, Col.debt, Col.CO2, Col.CPI)
Col.mean= ((Col$Col.debt*Col$Col.CO2*Col$Col.CPI)^(1/3))
Col = cbind(Col, Col.mean)
Col

Cos = cbind(Cos.debt.1, Cos.CO2) 
Cos = cbind(Cos, Cos.CPI.1) %>%
  dplyr::select(date, Cos.debt, Cos.CO2, Cos.CPI)
Cos.mean= ((Cos$Cos.debt*Cos$Cos.CO2*Cos$Cos.CPI)^(1/3))
Cos = cbind(Cos, Cos.mean)
Cos

Dom = cbind(Dom.debt.1, Dom.CO2) 
Dom = cbind(Dom, Dom.CPI.1) %>%
  dplyr::select(date, Dom.debt, Dom.CO2, Dom.CPI)
Dom.mean= ((Dom$Dom.debt*Dom$Dom.CO2*Dom$Dom.CPI)^(1/3))
Dom = cbind(Dom, Dom.mean)
Dom

Ecu = cbind(Ecu.debt.1, Ecu.CO2) 
Ecu = cbind(Ecu, Ecu.CPI.1) %>%
  dplyr::select(date, Ecu.debt, Ecu.CO2, Ecu.CPI)
Ecu.mean= ((Ecu$Ecu.debt*Ecu$Ecu.CO2*Ecu$Ecu.CPI)^(1/3))
Ecu = cbind(Ecu, Ecu.mean)
Ecu

Els = cbind(Els.debt.1, Els.CO2) 
Els = cbind(Els, Els.CPI.1) %>%
  dplyr::select(date, Els.debt, Els.CO2, Els.CPI)
Els.mean= ((Els$Els.debt*Els$Els.CO2*Els$Els.CPI)^(1/3))
Els = cbind(Els, Els.mean)
Els

Mex = cbind(Mex.debt.1, Mex.CO2) 
Mex = cbind(Mex, Mex.CPI.1) %>%
  dplyr::select(date, Mex.debt, Mex.CO2, Mex.CPI)
Mex.mean= ((Mex$Mex.debt*Mex$Mex.CO2*Mex$Mex.CPI)^(1/3))
Mex = cbind(Mex, Mex.mean)
Mex

Pan = cbind(Pan.debt.1, Pan.CO2) 
Pan = cbind(Pan, Pan.CPI.1) %>%
  dplyr::select(date, Pan.debt, Pan.CO2, Pan.CPI)
Pan.mean= ((Pan$Pan.debt*Pan$Pan.CO2*Pan$Pan.CPI)^(1/3))
Pan = cbind(Pan, Pan.mean)
Pan

Par = cbind(Par.debt.1, Par.CO2) 
Par = cbind(Par, Par.CPI.1) %>%
  dplyr::select(date, Par.debt, Par.CO2, Par.CPI)
Par.mean= ((Par$Par.debt*Par$Par.CO2*Par$Par.CPI)^(1/3))
Par = cbind(Par, Par.mean)
Par

Per = cbind(Per.debt.1, Per.CO2) 
Per = cbind(Per, Per.CPI.1) %>%
  dplyr::select(date, Per.debt, Per.CO2, Per.CPI)
Per.mean= ((Per$Per.debt*Per$Per.CO2*Per$Per.CPI)^(1/3))
Per = cbind(Per, Per.mean)
Per

Uru = cbind(Uru.debt.1, Uru.CO2) 
Uru = cbind(Uru, Uru.CPI.1) %>%
  dplyr::select(date, Uru.debt, Uru.CO2, Uru.CPI)
Uru.mean= ((Uru$Uru.debt*Uru$Uru.CO2*Uru$Uru.CPI)^(1/3))
Uru = cbind(Uru, Uru.mean)
Uru

Arg



countrynames = c("country", "date", "debt", "CO2", "CPI", "mean")

Argentina = "Argentina"
Argentina = rep(Argentina, length.out = 15)
Arg = cbind(Argentina, Arg)
colnames(Arg) = countrynames
Arg

Bolivia = "Bolivia"
Bolivia = rep(Bolivia, length.out = 15)
Bol = cbind(Bolivia, Bol)
Bol
colnames(Bol) = countrynames
Bol

Brazil = "Brazil"
Brazil = rep(Brazil, length.out = 15)
Bra = cbind(Brazil, Bra)
Bra
colnames(Bra) = countrynames
Bra

Chile = "Chile"
Chile = rep(Chile, length.out = 15)
Chi = cbind(Chile, Chi)
Chi
colnames(Chi) = countrynames
Chi

a = "Colombia"
b = rep(a, length.out = 15)
Col = cbind(b, Col)
colnames(Col) = countrynames

a = "Costa Rica"
b = rep(a, length.out = 15)
Cos = cbind(b, Cos)
colnames(Cos) = countrynames
Cos

a = "Dominican Republic"
b = rep(a, length.out = 15)
Dom = cbind(b, Dom)
colnames(Dom) = countrynames
Dom

a = "Ecuador"
b = rep(a, length.out = 15)
Ecu = cbind(b, Ecu)
colnames(Ecu) = countrynames
Ecu

a = "El Salvador"
b = rep(a, length.out = 15)
Els = cbind(b, Els)
colnames(Els) = countrynames
Els

a = "Mexico"
b = rep(a, length.out = 15)
Mex = cbind(b, Mex)
colnames(Mex) = countrynames
Mex

a = "Panama"
b = rep(a, length.out = 15)
Pan = cbind(b, Pan)
colnames(Pan) = countrynames
Pan

a = "Paraguay"
b = rep(a, length.out = 15)
Par = cbind(b, Par)
colnames(Par) = countrynames
Par

a = "Peru"
b = rep(a, length.out = 15)
Per = cbind(b, Per)
colnames(Per) = countrynames
Per

a = "Uruguay"
b = rep(a, length.out = 15)
Uru = cbind(b, Uru)
colnames(Uru) = countrynames
Uru

AB = rbind(Arg, Bol)  
ABr = rbind(AB, Bra)
ACh = rbind(ABr, Chi)
ACo = rbind(ACh, Col)
ACR = rbind(ACo, Cos)
AD = rbind(ACR, Dom)
AEc = rbind(AD, Ecu)
AEl = rbind(AEc, Els)
AM = rbind(AEl, Mex)
APn = rbind(AM, Pan)
APg = rbind(APn, Par)
APe = rbind(APg, Per)
AU = rbind(APe, Uru)
AU

df <- NULL  
df$country <- c("Argentina", "Bolivia", "Brazil", "Chile", "Costa Rica", "Ecuador", "Dominican Republic", "Ecuador", "El Salvador", "Mexico", "Panama", "Paraguay", "Peru", "Uruguay")
df$code<- c("ARG", "BOL", "BRA", "CHL", "COL", "CRI", "DOM", "ECU", "SLV", "MEX", "PAN", "PRY", "PER", "URY")  
df$mean = AU$mean
df<-as.data.frame(df)  
sPDF <- joinCountryData2Map( df, joinCode = "NAME", nameJoinColumn = "code")  

lat.am = mapCountryData(sPDF, nameColumnToPlot="mean", mapRegion='latin america',xlim=bbox(sPDFmyCountries)[1,], ylim=bbox(sPDFmyCountries)[2,])

sPDFmyCountries <- sPDF[sPDF$NAME %in% df$country,]
k = bbox(sPDFmyCountries)

get_map(location = k, maptype = "toner")
map = get_map(location = k, maptype = "toner")
```


```{r}
ggmap(map)
```

# Materials and methods

This section will describe data sources and coding methods. Before proceeding to the code and data itself however, it is important to discuss how this index will be defined.

Being that the Index is a model, it will be important to understand the key assumptions it
relies on, namely how it's inputs represent what they claim to represent. 

#### We can consider these casewise:

1. Political Stability - This will be represented by the level of public (or government)
debt in a single country. The thinking here is that as public debt levels increase, a 
nation's government will become increasingly unstable.

2. Economic Stability - One of the most common macroeconomic indicators of economic 
stability is the inflation rate. The Consumer Price Index of each country, a measure
tracking the prices of all goods in a country intended for end consumers, will be included. An unchecked inflation rate is always a danger to any economy.

3. Environmental Sustainability - The are a plethora of measure that may be used to track a nation's environmental sustainability. Although some of these measures may change over time, levels of Carbon Dioxide Emissions are always a concern. Therefore CO2 emissions by country will be included.


Thus the general workflow proceeds as follows:

* Gathering Requisite Data
* Constructing the Index
* Examine Trends



#### Here's a quick look at the packages used for this project

```{r, message=F, warning=F}
library(fredr)
library(wbstats)
library(dplyr)
library(ggplot2)
library(foreach)
library(spData)
library(sf)
library(plm)
library(forecast)
library(rworldmap)
library(ggmap)
```

### Download and clean data for Political Stability
```{r}
Arg = fredr_series_search_text("Argentina debt GDP")
head(Arg$title)
Arg$id[5]
Arg.debt = fredr("DDDM07ARA156NWDB")
dates = c(1999:2014)
dates.1 = c(1980:2017)
debtnames = c("date", "debt")
Arg.debt.1 = cbind(dates.1, Arg.debt) %>%
  filter(dates.1 >= 2000) %>%
  filter(dates.1 <= 2014) %>%
  dplyr::select(dates.1, value) %>%
  arrange((desc(dates.1)))
colnames(Arg.debt.1) <- c("date", "Arg.debt")

```

### Download and clean data for Economic Stability
```{r}
wbsearch(pattern = " CO2 emission")

CO2 = wb(indicator = "EN.ATM.CO2E.KT")

Arg.CO2 = CO2 %>%
  filter(country == "Argentina") %>%
  filter(date >= 2000) %>%
  dplyr::select(date, value) %>%
  arrange((desc(date)))
colnames(Arg.CO2) <- c("date1", "Arg.CO2")
```

### Download and clean data for Environmental Sustainability
```{r}
Arg.CPI = fredr_series_search_text("Argentina CPI")
head(Arg.CPI$title)
Arg.CPI$id[2]
Arg.CPI = fredr("DDOE01ARA086NWDB")
CPIdates.1 = c(1960:2014)
Arg.CPI.1 = cbind(CPIdates.1, Arg.CPI) %>%
  filter(CPIdates.1 >= 2000) %>%
  filter(CPIdates.1 <= 2014) %>%
  dplyr::select(CPIdates.1, value) %>%
  arrange((desc(CPIdates.1)))
colnames(Arg.CPI.1) <- c("date2", "Arg.CPI")
```

#### With all of the necessary data collected, it is now possible to construct the Development Index for each country 


```{r}
Arg = cbind(Arg.debt.1, Arg.CO2) 
Arg = cbind(Arg, Arg.CPI.1) %>%
  dplyr::select(date, Arg.debt, Arg.CO2, Arg.CPI)
Arg.mean= ((Arg$Arg.debt*Arg$Arg.CO2*Arg$Arg.CPI)^(1/3))
Arg = cbind(Arg, Arg.mean)
```
Note the object Arg.mean. This variable is the cubed root of a product of three inputs. To calculate this mean debt level, CO2 emisssions, and the CPI are all multiplied together. Because there are three inputs, the cubed root is taken. This is called a geometric mean, as opposed to the familiar arthimetic mean. The reason for using the geometric mean is that it is the same measure the UNDP uses to calculate the Human Development Index

#### Binding Countries by row yields the following table
```{r}
str(AU)
head(AU)
```


# Results




####First let's examine what the individual country level trends look like:
```{r}
ggplot(AU, aes(date, mean, col = country))+
  geom_point()+
  geom_line()+
  facet_wrap(~country)
```
Note that as the indicator goes up, utility goes down.

Next is it possible to identify a trend? 

There isn't enough data yet to build a time-series model of any one single country.
However, if we use a panel data model, it is possible to extend our analysis across borders.

##### Fixed Effects
```{r}
AU.fe <- plm(mean~date, data = AU, model = "within")
summary(AU.fe)
```
Using a Fixed Effects model explains just under 30% of the variation we see in our model

##### Random Effects
```{r}
AU.ra <- plm(mean~date, data = AU, model = "random")
summary(AU.ra)
```
Using a Random Effects model explains just over 30% of the variation we see in our model




# Conclusions

Indices are an interesting may of measuring trends over time and space. There are many that exist that provide us with new information every year. Creating a new index may not be too difficult conceptually, but there are challenges to building one that is effective. There should be a strong case that a combination of quantitative variables represent a changing process in the world. Next, ensuring a sufficient amount of data exists is necessary for building the strongest time-series or panel data models.
